How demand pattern identification and multicriteria ABC classification can guide management decisions

### 2023-12-26 manuscript/FCPCOM/manuscript.tex
A study on multicriteria ABC and demand pattern classifications

\title{A study on multicriteria ABC classification with analytical hierarchy process}

@@

\begin{abstract} 
Multicriteria classification is usualy very important to the decision-making in manufacturing management process. For such classification, the attribution of weights to the criteria strongly influences the coherence of the results found. Saaty's Analytic Hierarchy Process (AHP) is an important method for assigning weights to multiple criteria.  AHP's logic is not complicated at all but, since matrices of pairwise comparisons of criteria  are usually generated manually and based only on some employee know-how, there is a huge complexity on generating a consistent pairwise matrix. Especially when many criteria are used. This paper presents a constructive algorithm that can be used to adjust inconsistent matrices, forcing such matrices to have a better consistency rate. We tested this algorithm by applying the AHP method, for multicriteria ABC classification, to companies in two sectors. As a result we observed that the algorithm can adjust the pairwise matrices in just a few seconds, avoiding the manual work that would be done in weeks, therefore showing that it is an important resource for applying the AHP method. We also present in this paper an analysis of the importance of the attribution of the weights to the criteria and show how the multicriteria classification may influence the decision on the choice of the appropriate forecasting method.
\end{abstract}
%%Graphical abstract
%\begin{graphicalabstract}
%\end{graphicalabstract}

%%Research highlights
\begin{highlights}
\item generative method to force consistency of pairwise comparisons matrix;
\item importance of the correct balance of product criteria weights assigning;
\item importance of ABC multicriteria classification for forecasting.
\end{highlights}

\begin{keyword}
multicriteria classification \sep ABC classification \sep analytic hierarchy process \sep pairwise matrix consistency \sep COPSolver
\end{keyword}
\end{frontmatter}

@@

The algorithm is not represented correctly.

Changes will be made during the preparation of the article according to the module of software COPSolver: library for solving classification problems.

\begin{algorithm}
\caption{function $consistencyRate()$}\label{alg:CA}
\begin{algorithmic}
\Require $[a_{ij}]_{i,j=1}^{n}, RI[x]_{x=1}^{10}$
\State{$\lambda_{max} = mainEigenvalue([a_{ij}]_{i,j=1}^{n})$ \footnote{main eigenvalue can be found using "Eigen/Eigenvalues library.}}
\State{$CI = (\lambda_{max} - n) / (n-1)$}
\State{$CR \gets CI/RI[n]$}
\State{\Return {$CR$}}	
\end{algorithmic}
\end{algorithm}

\subsection{Constructive algorithm}

\begin{algorithm}
\caption{function $constructivelyForceConsistency([a_{ij}]_{i,j=1}^{n})$}\label{alg:CA}
\begin{algorithmic}
\Require $[a_{ij}]_{i,j=1}^{n}$
\Ensure $consistencyRate([a_{ij}]_{i,j=1}^{n}) \leq 0.1$
\State{$CR = consistencyRate([a_{ij}]_{i,j=1}^{n})$}
\If {$CR \leq 0.1 $}
	\State{\Return {$[a_{ij}]_{i,j=1}^{n}$}}
\Else
	\For{$ 3 \leq k \leq n $}
		\State{$([a_{ij}]_{i,j=1}^{k}, CR) \gets forceConsistency([a_{ij}]_{i,j=1}^{k})$}
		\If{$CR \leq 0.1$}
			\State{\textbf{break for}}
		\EndIf
	\EndFor
\EndIf
\State{\Return {$([a_{ij}]_{i,j=1}^{n}, CR)$}}

\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{function $forceConsistency([a_{ij}]_{i,j=1}^{k})$}\label{alg:CA}
\begin{algorithmic}
\Require $[a_{ij}]_{i,j=1}^{n} \quad | \quad consistencyRate([a_{ij}]_{i,j=1}^{n-1}) \leq 0.1 $
\Ensure $consistencyRate([a_{ij}]_{i,j=1}^{n}) \leq 0.1$
\State{$CR \gets consistencyRate([a_{ij}]_{i,j=1}^{n})$}
\For{$j \leq n-2$}
	\For{$j+1 \leq k \leq n-1 $}
			\State{$g \gets 0$}
			\State{$s \gets 0$}
		\If{$a_{nj} > a_{nk} \textrm{ and } a_{jk} \leq 1 $}
			\State{$g \gets j$}
			\State{$s \gets k$}
		\ElsIf{$a_{nj} < a_{nk} \textrm{ and } a_{jk} \geq 1 $}
			\State{$g \gets k$}
			\State{$s \gets j$}
		\EndIf
		\If{$g \neq 0$}
			\State{$([a_{ij}]_{i,j=1}^{n} , CR) \gets reduce([a_{ij}]_{i,j=1}^{n}, g, s)$}
			\If{$CR \geq 0.1$}
				\State{$([a_{ij}]_{i,j=1}^{n} , CR) \gets encrease([a_{ij}]_{i,j=1}^{n}, g, s)$}
			\EndIf
		\EndIf
		\If{$CR \leq 0.1$}
			\State{\textbf{break for}}
		\EndIf
	\EndFor
	\If{$CR \leq 0.1$}
		\State{\textbf{break for}}
	\EndIf
\EndFor
\State{\Return {$([a_{ij}]_{i,j=1}^{n}, CR)$}}
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{function $reduce([a_{ij}]_{i,j=1}^{n}, g, s)$}\label{alg:CA}
\begin{algorithmic}
\Require $g,s < n; [a_{ij}]_{i,j=1}^{n}$
\While{$a_{ng} > a_{ns}$}
	\State{$a_{ng} \gets a_{ng}^{--}$}
	\State{$a_{gn} \gets a_{gn}^{++}$}
	\State{$aux_{CR} = consistencyRate([a_{ij}]_{i,j=1}^{n})$}
	\If{$aux_{CR} < CR$}
		\State{$CR = aux_{CR}$}
		\If{$CR \leq 0.1$}
			\State{\textbf{break}}
		\EndIf
	\Else
		\State{$a_{ng} \gets a_{ng}^{++}$}
		\State{$a_{gn} \gets a_{gn}^{--}$}
		\State{\textbf{break}}
	\EndIf
\EndWhile
\State{\Return {$([a_{ij}]_{i,j=1}^{n}, CR)$}}
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{function $encrease([a_{ij}]_{i,j=1}^{n}, g, s)$}\label{alg:CA}
\begin{algorithmic}
\Require $g,s < n; [a_{ij}]_{i,j=1}^{n}$
\While{$a_{ng} > a_{ns}$}
	\State{$a_{ns} \gets a_{ns}^{++}$}
	\State{$a_{sn} \gets a_{sn}^{--}$}
	\State{$aux_{CR} = consistencyRate([a_{ij}]_{i,j=1}^{n})$}
	\If{$aux_{CR} < CR$}
		\State{$CR = aux_{CR}$}
		\If{$CR \leq 0.1$}
			\State{\textbf{break}}
		\EndIf
	\Else
		\State{$a_{ns} \gets a_{ns}^{--}$}
		\State{$a_{sn} \gets a_{sn}^{++}$}
		\State{\textbf{break}}
	\EndIf
\EndWhile
\State{\Return {$([b_{ij}]_{i,j=1}^{n}, CR)$}}
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{function $a_{ij}^{++}$}\label{alg:CA}
\begin{algorithmic}
\Require $a_{ij} \quad | \quad a_{ij} \in \{1/9, 1/7, 1/5, 1/3, 1, 3, 5, 7, 9\}$
	\If{$a_{ij} \geq 1 \textrm{ and } a_{ij} \neq 9 $}
		\State {$a_{ij} \gets a_{ij} + 2$}
	\ElsIf{$a_{ij} < 1$}
		\State {$a_{ij} \gets 1/(1/a_{ij} - 2)$}
	\EndIf
\State{\Return {$a_{ij}$}}
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
\caption{function $a_{ij}^{--}$}\label{alg:CA}
\begin{algorithmic}
\Require $a_{ij} \quad | \quad a_{ij} \in \{1/9, 1/7, 1/5, 1/3, 1, 3, 5, 7, 9\}$
	\If{$a_{ij} \geq 3  $}
		\State {$a_{ij} \gets b_{ij} - 2$}
	\ElsIf{$a_{ij} \leq 1 \textrm{ and } a_{ij} \neq 1/9$}
		\State {$a_{ij} \gets 1/(1/a_{ij} + 2)$}
	\EndIf
\State{\Return {$a_{ij}$}}
\end{algorithmic}
\end{algorithm}

@@

>>>>

\title{How demand pattern identification and multicriteria ABC classification can guide management decisions}

@@

\begin{abstract} 
Multicriteria classification is usualy very important to the decision-making in manufacturing management process. For such classification, the attribution of weights to the criteria strongly influences the coherence of the results found. Saaty's Analytic Hierarchy Process (AHP) is an important method for assigning weights to multiple criteria.  AHP's logic is not complicated at all but, since matrices of pairwise comparisons of criteria  are usually generated manually and based only on some employee know-how, there is a huge complexity on generating a consistent pairwise matrix. Especially when many criteria are used. This paper presents a constructive algorithm that can be used to adjust inconsistent matrices, forcing such matrices to have a better consistency rate. We tested this algorithm by applying the AHP method, for multicriteria ABC classification, to companies in two sectors. As a result we observed that the algorithm can adjust the pairwise matrices in just a few seconds, avoiding the manual work that would be done in weeks, therefore showing that it is an important resource for applying the AHP method. We also present in this paper an analysis of the importance of the attribution of the weights to the criteria and show how the multicriteria and demand pattern classifications may influence the decision on the choice of the appropriate forecasting method.
\end{abstract}
%%Graphical abstract
%\begin{graphicalabstract}
%\end{graphicalabstract}

%%Research highlights
\begin{highlights}
\item literature review about multicriteria classification;
\item literature review about demand pattern classification;
\item generative method to force consistency of pairwise comparisons matrix;
\item importance of the correct balance of product criteria weights assigning;
\item importance of ABC multicriteria and demand pattern classifications for forecasting;
\item brief description of COPSolver: library for solving the multicriteria classification problem;
\item brief description of COPSolver: library for solving the demand pattern classification problem.
\end{highlights}

\begin{keyword}
demand pattern identitification \sep multicriteria ABC classification \sep analytic hierarchy process \sep pairwise matrix consistency \sep COPSolver
\end{keyword}
\end{frontmatter}

@@

@@

To test the two developed COPSolver libraries (COPSolver: library for solving the multi-criteria classification problem and COPSolver: library for solving the demand pattern classification problem), we used data from three companies in three different sectors (plastic packaging manufacturers, furniture retailers and car mechanics). In the case of the company in the car mechanics sector, only the data relating to truck repairs was used; in the case of the other two companies, all the data obtained from all the products sold over the last 5 years was used. The files containing the formatted data used for all the tests can be found at tbfraga.github.io/COPSolver/benchmarks.
